---
title: "2025-05-20"
format: html
editor: visual
---

```{r}

water_quality <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-20/water_quality.csv')
weather <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-20/weather.csv')
```
```{r}
library(tidyverse)
glimpse(water_quality)
glimpse(weather)
```
# Research question

Does daily rainfall influence bacterial contamination (enterococci) at Sydney swim sites?


```{r}
df_site <- water_quality %>%
  filter(!is.na(enterococci_cfu_100ml)) %>%
  group_by(region, date) %>%
  summarize(
    avg_entero = mean(enterococci_cfu_100ml, na.rm = TRUE),
    avg_temp   = mean(water_temperature_c,    na.rm = TRUE),
    avg_cond   = mean(conductivity_ms_cm,     na.rm = TRUE),
    n_samples  = n(),
    .groups     = "drop"
  ) %>%
  left_join(weather %>% select(date, precipitation_mm), by = "date") %>%
  filter(!is.na(precipitation_mm))

# Pick top 4 regions by sample count
top_regions <- df_site %>%
  count(region, sort = TRUE) %>%
  slice_head(n = 4) %>%
  pull(region)

df_top <- df_site %>% filter(region %in% top_regions)

# Plot
ggplot(df_top, aes(
    x = precipitation_mm,
    y = avg_entero,
    colour = avg_temp,
    size = avg_cond
  )) +
  geom_point(alpha = 0.7) +
  scale_y_log10(
    name   = "Avg Enterococci (CFU/100 mL)\n(log scale)",
    breaks = c(10, 100, 500, 1000)
  ) +
  scale_x_continuous(name = "Daily Precipitation (mm)") +
  scale_colour_viridis_c(
    option = "plasma",
    name   = "Water Temp (°C)"
  ) +
  scale_size_continuous(
    range = c(1, 6),
    name  = "Conductivity\n(mS/cm)"
  ) +
  geom_smooth(
    method    = "lm",
    aes(weight = n_samples),
    se        = FALSE,
    colour    = "black",
    linetype  = "dashed",
    linewidth = 0.5
  ) +
  facet_wrap(~ region, scales = "free") +
  labs(
    title    = "Rainfall, Bacteria & Water Conditions at Sydney Swim Sites",
    subtitle = "Site-day means: enterococci (log), coloured by temp, sized by conductivity",
    caption  = "TidyTuesday 2025-05-20 | github.com /tidytuesday"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "right",
    plot.subtitle   = element_text(margin = margin(b = 8)),
    plot.caption    = element_text(size = 8, color = "gray40")
  )
```


**What the plot shows:** Across all four Sydney regions, heavier rain days generally correspond to higher average enterococci (upward‐sloping fits), with warmer water and greater conductivity often amplifying contamination; the strength of these associations varies by region.

**Key caveats:**

* **Precipitation** comes from one station but applies to many sites, so local downpours may be missed.
* **Site‐day averages** mask extreme values at individual beaches.
* **Sampling times** vary, so rain–sample timing mismatches can blur true rain impacts.


